---
title: "GenmStats_Week3"
author: "Alisa"
date: "2/26/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Week 3 Quiz: 

1. Logistic modeling
2. Study statistics

##Question 1: Fit a linear model and a logistic regression model to the data for the 3rd SNP. What are the coefficients for the SNP variable? How are they interpreted? (Hint: Don't forget to recode the 0 values to NA for the SNP data)

1. Retrieve SNP data, isolate snp[,3]
2. Recode 0 values as NA
3. Linear model using lm
4. Log model using glm

```{r Question 1}
library(snpStats)
library(broom)
data(for.exercise)
use <- seq(1, ncol(snps.10), 10)
sub.10 <- snps.10[,use]
snpdata = sub.10@.Data
status = subject.support$cc
snp3 = as.numeric(snpdata[,3])
snp3[snp3==0] = NA
lm1 <-  lm(status ~ snp3)
glm1 = glm(status ~ snp3,family="binomial")
plot(status, col = as.numeric((snp3)))
```

## Question 3: Fit a logistic regression model on a recessive (need 2 copies of minor allele to confer risk) and additive scale for the 10th SNP. Make a table of the fitted values versus the case/control status. Does one model fit better than the other?

1. Recessive scale: snp10 == 2
2. Additive scale as done above

```{r Question 3}
snpdata = sub.10@.Data
snp10 = as.numeric(snpdata[,10])
snp10[snp10==0] = NA
snp10_rec = (snp10 == 2)
glm_rec <- glm(status ~ snp10_rec,family="binomial")
glm2 = glm(status ~ snp10,family="binomial")
table(status, glm_rec$fitted.values)
table(status, glm2$fitted.values)
```

## Question 4: Fit an additive logistic regression model to each SNP. What is the average effect size? What is the max? What is the minimum?

1. Use snp.rhs.tests to do logistic regression on each snp of snp data
2. 

```{r Question 4}
results = rep(NA, dim(snpdata)[2])
for (i in 1:ncol(snpdata)){  # this line opens the loop and defines how often it will be executed
  snpdata_i = as.numeric(snpdata[,i])
  snpdata_i[snpdata_i == 0] = NA
  glm_i = glm(status ~ snpdata_i, family = "binomial")
  results[i] = tidy(glm_i)$statistic[2]# In the final loop code, save the results into the dummy vector
}                                   # close the loop

mean(results)
min(results)
max(results)
```
##Question 5: Fit an additive logistic regression model to each SNP and square the coefficients. What is the correlation with the results from using ðšœðš—ðš™.ðš›ðš‘ðšœ.ðšðšŽðšœðšðšœ and ðšŒðš‘ðš’.ðšœðššðšžðšŠðš›ðšŽðš? Why does this make sense?

```{r Question 5}
#results_coeff = rep(NA, dim(snpdata)[2]) # set up a dummy vector to save all values
#for (i in 1:ncol(snpdata)){         # this line opens the loop and defines how often it will be executed
  #snpdata_i = as.numeric(snpdata[,i])
  #snpdata_i[snpdata_i == 0] = NA
  #glm_i = glm(status ~ snpdata_i, family = "binomial")
  #results_coeff[i] = glm_i$coefficients # In the final loop code, save the results into the dummy vector
}                                   # close the loop
#results_coeff_squre =  results_coeff^2
results_sqre = results^2
glm_all = snp.rhs.tests(status ~ 1,snp.data=sub.10)
qq.chisq(chi.squared(glm_all),df=1)
cor(results_sqre, chi.squared(glm_all))
```

##Question 6: Do the log2(data + 1) transform and fit calculate F-statistics for the difference between studies/populations using genefilter:rowFtests and using genefilter:rowttests. Do you get the same statistic? Do you get the same p-value?

```{r Question 6}
#edata_mp_log
library(genefilter)
fstats_obj = rowFtests(edata,as.factor(pdata$lane.number))
```
##Question 7: First test for differences between the studies using the ð™³ð™´ðš‚ðšŽðššðŸ¸ package using the ð™³ð™´ðš‚ðšŽðšš function. Then do the log2(data + 1) transform and do the test for differences between studies using the ðš•ðš’ðš–ðš–ðšŠ package and the ðš•ðš–ð™µðš’ðš, ðšŽðš‹ðšŠðš¢ðšŽðšœ and ðšðš˜ðš™ðšƒðšŠðš‹ðš•ðšŽ functions. What is the correlation in the statistics between the two analyses? Are there more differences for the large statistics or the small statistics (hint: Make an MA-plot).

```{r Question 7}
library(DESeq2)
library(limma)
de = DESeqDataSetFromMatrix(edata_mp, pdata_mp, ~study)
glm_all_nb = DESeq(de)
result_nb = results(glm_all_nb)
mod = model.matrix(~ pdata_mp$study)
fit_limma = lmFit(edata_mp_log,mod)
ebayes_limma = eBayes(fit_limma)
head(ebayes_limma$t)
limma_all = topTable(ebayes_limma, number = dim(edata_mp_log)[1])
plotMA(result_nb, limma_all)
```

##Question 8: 

```{r Question 8}
#number with significant false discovery rate at 0.05 for DESeq
fp_bh = p.adjust(result_nb$pvalue, method = "BH")
hist(fp_bh, col = 3)
quantile(fp_bh)
sum(fp_bh < 0.05)

#number with significant false discovery rate at 0.05 for limma
fp_bh_limma = p.adjust(limma_all$P.Value, method = "BH")
hist(fp_bh_limma, col = 4)
quantile(fp_bh_limma)
sum(fp_bh_limma < 0.05)